SDG contributions

### 2023-09-15 article/article.qmd
conclude baseline calculations for world regions

subset_countries <- filter(
  mag_2021_normalised_region_share_time, 
  sub_continent %in% c("Northern America", "Western Europe", "Northern Europe",
                       "Eastern Asia"
  ))

mag_2021_normalised_region_share_time %>% 
  ggplot(aes(as_year(year), share, group = sub_continent, colour = sub_continent)) +
  geom_line(colour = base_col) +
  geom_line(data = subset_countries) +
  geom_point(data = subset_countries, size = 1.2) +
  date_scale +
  scale_y_continuous(labels = function(x) scales::percent(x, 1)) +
  colorspace::scale_color_discrete_qualitative() +
  theme(legend.position = "top") +
  labs(x = NULL, y = "% of publications from world region across MAG",
       colour = NULL)
```

The above is the same graph above but for all of MAG. The next step is to 
integrate this into the existing graph. Options:

- Normalise existing graph by baseline
- Add baseline as a background (maybe with different line type)

The latter seems more in line with the focus of the analysis, since the target
is the overall distribution of knowledge production, not the relative 
specialisation.

>>>>

mag_2021_subset_countries <- filter(
  mag_2021_normalised_region_share_time, 
  sub_continent %in% c("Northern America", "Western Europe", "Northern Europe",
                       "Eastern Asia"
  ))

mag_2021_normalised_region_share_time %>% 
  ggplot(aes(as_year(year), share, group = sub_continent, colour = sub_continent)) +
  geom_line(colour = base_col) +
  geom_line(data = mag_2021_subset_countries) +
  geom_point(data = mag_2021_subset_countries, size = 1.2) +
  date_scale +
  scale_y_continuous(labels = function(x) scales::percent(x, 1)) +
  colorspace::scale_color_discrete_qualitative() +
  theme(legend.position = "top") +
  labs(x = NULL, y = "% of publications from world region across MAG",
       colour = NULL)
```

## Joint figure including baseline
```{r, fig.width=8.2, fig.height=4.4}
#| fig-cap: "Proportion of research from world regions on SDGs. The dashed line
#|           indicates the average across all of MAG"
#| label: fig-sub_continents_over_time_with_baseline
normalised_region_share_time %>% 
  ggplot(aes(as_year(year), share, group = sub_continent, colour = sub_continent)) +
  geom_line(colour = base_col) +
  geom_line(data = mag_2021_subset_countries, linetype = 2) +
  geom_line(data = subset_countries) +
  geom_point(data = subset_countries, size = 1.2) +
  facet_wrap(vars(fix_sdg(SDG_label))) +
  date_scale +
  scale_y_continuous(labels = function(x) scales::percent(x, 1)) +
  colorspace::scale_color_discrete_qualitative() +
  theme(legend.position = "top") +
  labs(x = NULL, y = "% of publications from world region within SDGs",
       colour = NULL)
```

```{r}
# look at relative specialisation to identify how global south is represented in
# SDG 2

continent_specialisation <- normalised_region_share_time %>% 
  rename(nn_sdg = nn, share_sdg = share) %>% 
  mutate(year = as.character(year)) %>% 
  left_join(mag_2021_normalised_region_share_time) %>% 
  mutate(specialisation = share_sdg / share)
```

```{r, fig.width=10, fig.height=10}
#| fig-cap: "Specialisation of world regions. The specialisation factor is
#| calulated by dividing the share of research towards a certain SDG by the 
#| overall share of research coming from the same region. Melanesia, Polynesia
#| and Micronesia are omitted due to low cell counts."
continent_specialisation %>% 
  filter(!sub_continent %in% c("Melanesia", "Polynesia", "Micronesia")) %>% 
  ggplot(aes(as_year(year), specialisation, colour = fix_sdg(SDG_label))) +
  geom_hline(yintercept = 1, colour = "grey30") +
  geom_line() +
  geom_point(size = 1.1) +
  facet_wrap(vars(sub_continent)) +
  colorspace::scale_color_discrete_qualitative() +
  theme(legend.position = "top") +
  labs(x = NULL, colour = NULL, y = "Specialisation factor")
```
